---
output:
  html_document:
    includes:
      before_body: "header.html"
    css: "styles.css"
    toc: true
    toc_depth: 2
    toc_float: 
      smooth_scroll: true
    number_sections: true
    highlight: tango
    # code_folding: hide   # This collapses the code chunks by default
---

# Introduction

This report presents a **Principal Component Analysis (PCA)** of
employment data from Eurostat. The goal is to:

1.  **Identify patterns and relationships** between countries based on
    employment statistics for different citizen categories.
2.  **Reduce dimensionality** for easier visualization and
    interpretation.
3.  **Determine the contributions** of each variable (citizen category)
    to the principal components.
4.  **Visualize countries' similarities and differences** in employment
    patterns.

------------------------------------------------------------------------

# Loading necessary Libraries

Loading the necessary libraries for data manipulation, analysis, and
visualization.

```{r load-libraries, message=FALSE, warning=FALSE}
# Load necessary libraries
library(tidyverse)     
library(eurostat)      
library(FactoMineR)   
library(factoextra)   
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyverse)
library(lubridate)
library(plotly)
library(scales)  
library(patchwork) 
library(gridExtra)  
library(highcharter)
library(knitr)
library(kableExtra)
library(htmltools)
```

------------------------------------------------------------------------

# Data Import and Exploration

The `lfsa_egan` dataset was obtained from [EUROSTAT’s official website](https://ec.europa.eu/eurostat/databrowser/explore/all/popul?lang=en&subtheme=labour.employ.lfsa.lfsa_emp&display=list&sort=category&extractionId=lfsa_egan). This dataset belongs to the *Labour Force Survey (LFS)* and contains annual employment data categorized by different demographics and economic dimensions. <br>
- **Dataset**: 496671 observations and 8 variables. <br>
- **Types of variables**: it mainly contains categorical variables and one numerical variable.

```{r setup, message=FALSE, warning=FALSE, class.source = 'fold-show'}
data <- get_eurostat("lfsa_egan")
```

In order to have a better understanding of the dataset, a brief description for each variable and the corresponding unique values are provided.

```{r echo=TRUE}
# Sample data frame
df <- data.frame(
  Variable = c(
    "TIME_PERIOD", "geo", "citizen", "age", "unit",
    "sex", "freq", "values"
  ),
  Description = c(
    "Sampling year",
    "Geopolitical entity",
    "Country of citizienship",
    "Age ranges",
    "Measurement's unit for employment",
    "Gender",
    "Sampling frequency",
    "Values"
  ),
  Type = c("date", "chr", "chr", "chr", "chr", "chr", "chr", "num"),
  Unique_values = c(
    paste0("`", unique(data$TIME_PERIOD), "`", collapse = " "),
    paste0("`", unique(data$geo), "`", collapse = " "), 
    paste0("`", unique(data$citizen), "`", collapse = " "), 
    paste0("`", unique(data$age), "`", collapse = " "), 
    "`THS_PER` (Thousand persons)", 
    "`F` (Female) `M` (Male) `T` (Total)", 
    "`A` (Annual)", 
    "`numerical values"
  )
)

# Create and style the table
kable(df, "html", escape = FALSE, col.names = c("Variable", "Description", "Type", "Unique_values")) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover"),
    font_size = 14
  ) %>%
  row_spec(0, bold = TRUE) %>%  # Make the header row bold
  column_spec(1, bold = TRUE) %>%   # Make the first column (Variable) bold
  scroll_box(width = "100%", height = "100%", fixed_thead	= FALSE)

```

------------------------------------------------------------------------

# Trends over time

To better understand how employment levels for different citizenship
groups have changed over time (1995-2023) the following analysis was
conducted, helping to identify noticeable trends or anomalies during the
time period. The dataset was filtered, leading to a subset containing
all observations for the **Working Age Population** (`Y15-64`) and
including undifferentiated data for Females and Males.

```{r, message=FALSE, warning=FALSE, out.width='100%'}
# Filter data for working-age population (15-64) and 'Total' for sex
working_age_data <- data %>%
  filter(age == "Y15-64", sex == "T") %>%
  mutate(
    year = year(TIME_PERIOD),
    employment_real = values * 1000  # Convert employment values to actual numbers
  )

# Group employment by year and citizenship
time_trend <- working_age_data %>%
  group_by(TIME_PERIOD, citizen) %>%
  summarise(total_employment = sum(employment_real, na.rm = TRUE)) %>%
  ungroup()

# Create the Highcharter plot
hchart <- highchart() %>%
  hc_chart(type = "line") %>%
  hc_title(text = "Employment Trends by Citizenship Group (1995-2023)") %>%
  hc_subtitle(text = "Working Age: 15-64 | Females and Males together") %>%
  hc_xAxis(
    type = "datetime",
    title = list(text = "Year"),
    labels = list(format = "{value:%Y}")
  ) %>%
  hc_yAxis(
    title = list(text = "Total Employment"),
    labels = list(format = "{value:,}")
  ) %>%
  hc_tooltip(
    shared = TRUE,
    crosshairs = TRUE,
    pointFormat = "<b>Citizen Group:</b> {series.name}<br><b>Total Employment:</b> {point.y:,.0f}<br>"
  ) %>%
  hc_colors(c(
    "#003399",  # EU27_2020_FOR (Dark Blue)
    "#DAA520",  # FOR (Goldenrod)
    "#228B22",  # NAT (Forest Green)
    "#00CED1",  # NEU27_2020_FOR (Dark Turquoise)
    "#1E90FF",  # NRP (Dodger Blue)
    "#9370DB",  # STLS (Medium Purple)
    "#FF1493"   # TOTAL (Deep Pink)
  )) %>%
  hc_plotOptions(
    series = list(marker = list(enabled = FALSE))
  )

# Add series for each citizenship group
citizenship_groups <- unique(time_trend$citizen)

for (citizen in citizenship_groups) {
  citizen_data <- time_trend %>% filter(citizen == !!citizen)
  
  hchart <- hchart %>%
    hc_add_series(
      data = list_parse2(data.frame(
        x = as.numeric(as.POSIXct(citizen_data$TIME_PERIOD)) * 1000,
        y = citizen_data$total_employment
      )),
      name = citizen
    )
}

# Define the event annotations
event_annotations <- data.frame(
  date = as.Date(c("2001-01-01", "2004-01-01", "2007-01-01", "2008-01-01", "2020-01-01"))
  )

# Add vertical dashed lines for each event annotation with hover tooltips
plot_lines <- lapply(1:nrow(event_annotations), function(i) {
  list(
    color = "#084594",          # Dark blue for the line
    width = 2,                  # Line thickness
    value = as.numeric(as.POSIXct(event_annotations$date[i])) * 1000,
    dashStyle = "Dash",         # Dashed line style
    zIndex = 3,                 # Ensure lines are above the plot
    label = list(
      text = "",                # No static text
      style = list(fontSize = "0px")  # Hide the label
    )
  )
})

# Add the plot lines to the Highcharter plot
hchart <- hchart %>%
  hc_xAxis(plotLines = plot_lines)

# Enable exporting and display the plot
hchart <- hchart %>%
  hc_exporting(enabled = TRUE) %>%
  hc_legend(enabled = TRUE)

# Display the chart
hchart
```

The dotted-blue lines represent relevant events that might have
influenced the employment trend:

-   **2001**: The combination of Euro adoption preparations, economic
    growth, labour reforms, and anticipation of EU enlargement
    contributed to the significant rise in employment levels recorded
    in 2001. This was a period of optimism and structural change that
    positively impacted employment across various sectors in the EU.
-   **2004**: EU Enlargement saw a significant rise in employment
    levels, particularly for native citizens and total employment.
-   **2007**: Bulgaria & Romania Join the EU led to a continued increase
    in employment levels.
-   **2008**: Financial Crisis led to a sharp decline in employment
    trends due to widespread business failures, reduced investments, and
    economic contraction, resulting in significant job losses across
    many sectors in the EU
-   **2020**: COVID-19 Pandemic caused a slight dip in employment,
    reflecting the economic disruption during this period.

### Overall Employment Growth:

Total Employment (`TOTAL`) shows a sharp increase between 2001 and 2007,
stabilizing afterward. This increase coincides with the events following
the Euro adoption preparations (1999-2002), the 2004 EU Enlargement and
2007 accession of Bulgaria and Romania.

### Native Employment:

Employment of Native Citizens (`NAT`) follows a similar pattern to the
total employment trend, meaning that they represent the majority of the
employment. The growth stabilizes post-2007 and experiences a slight dip
during the COVID-19 pandemic in 2020.

### EU citizens working in other EU countries:

Employment of EU citizens working in other EU countries
(`EU27_2020_FOR`) shows a relatively stable trend with moderate growth
over the period.

### Non-EU foreign citizens:

Employment of non-EU foreign citizens (`FOR`) exhibits a gradual but
consistent increase, indicating the growing role of foreign labor in the
EU economy.

### No Response:

No Response (`NRP`) category shows a drop after 2003, possibly showing
the positive consequences of the 2004 EU Enlargement.

------------------------------------------------------------------------

# Gender-Based Employment Trends

Two interactive plots were displayed to present employment levels by
Gender over Time and across Countries.

## Employment Levels by Gender for the Working Age Population (15-64) Over Time

The following plot indicates the employment levels by Gender aged 15-64
across all countries, excluding EU-wide aggregates (EU27_2020 and EA20).
The data spans multiple years and provides insights into employment
trends by gender.

```{r, message=FALSE, warning=FALSE, out.width='100%'}
# Filter the data for working age (Y15-64), Males (M), Females (F), and exclude EU27_2020 and EA20
  filtered_data <- data %>%
    filter(
      age == "Y15-64",
      sex %in% c("M", "F"),
      !geo %in% c("EU27_2020", "EA20")
    ) %>%
    mutate(
      year = year(TIME_PERIOD),
      employment_real = values * 1000  # Convert employment values to actual numbers
    )
  
  # Summarize employment by year and sex
  summarized_data <- filtered_data %>%
    group_by(year, sex) %>%
    summarise(total_employment = sum(employment_real, na.rm = TRUE)) %>%
    ungroup()
  
  # Create the Highcharter bar plot
  hchart <- highchart() %>%
    hc_chart(type = "column") %>%
    hc_title(text = "Employment Levels by Gender for the Working Age Population (15-64) Over Time") %>%
    hc_subtitle(text = "All countries (excluding EU aggregates EU27_2020 - EA20)") %>%
    hc_xAxis(
      categories = unique(summarized_data$year),
      title = list(text = "Year"),
      labels = list(rotation = 45)  # Rotate x-axis labels for readability
    ) %>%
    hc_yAxis(
      title = list(text = "Total Employment"),
      labels = list(format = "{value:,}")  # Format y-axis with commas
    ) %>%
    hc_plotOptions(column = list(grouping = TRUE)) %>%
    hc_tooltip(shared = TRUE, pointFormat = "<b>{series.name}</b>: {point.y:,.0f}<br>") %>%
    hc_add_series(
      data = summarized_data %>% filter(sex == "M") %>% pull(total_employment),
      name = "Male",
      color = "#003399"
    ) %>%
    hc_add_series(
      data = summarized_data %>% filter(sex == "F") %>% pull(total_employment),
      name = "Female",
      color = "#FFCC00"
    ) %>%
    hc_legend(enabled = TRUE) %>%
    hc_exporting(enabled = TRUE)  # Enable export options
  
  # Display the chart
  hchart
```

This visualization helps to identify patterns and trends in employment
by gender over the years, highlighting differences and potential shifts
in labor market participation for males and females. **Insert Comment
here**

## Employment Levels by Gender and Country (Working Age: 15-64)

The following interactive plot allows users to select different years to
observe how employment levels have changed over time across countries
(including the aggregates `EA20` and `EU27_2020`) for the working
popuation (15-64), providing insights into gender-specific employment
trends and patterns.

```{r, message=FALSE, warning=FALSE, out.width='100%'}

# Filter the data for working age (Y15-64), Males (M), Females (F), and citizen TOTAL
filtered_data <- data %>%
  filter(
    age == "Y15-64",
    sex %in% c("M", "F"),
    citizen == "TOTAL"
  ) %>%
  mutate(
    year = year(TIME_PERIOD),
    employment_real = values * 1000  # Convert employment values to actual numbers
  )

# Summarize employment by year, country, and sex
summarized_data <- filtered_data %>%
  group_by(year, geo, sex) %>%
  summarise(total_employment = sum(employment_real, na.rm = TRUE)) %>%
  ungroup()

# Ensure all combinations of year, geo, and sex exist, filling missing ones with zero
summarized_data <- summarized_data %>%
  complete(year, geo, sex, fill = list(total_employment = 0))

# Define the correct order of countries with EA20 and EU27_2020 first
geo_levels <- c("EA20", "EU27_2020", setdiff(unique(summarized_data$geo), c("EA20", "EU27_2020")))

# Apply the factor order to geo
summarized_data$geo <- factor(summarized_data$geo, levels = geo_levels)

# Create a highcharter plot with a dropdown for year selection
years <- unique(summarized_data$year)

# Initialize highchart
hchart <- highchart() %>%
  hc_chart(type = "column") %>%
  hc_title(text = "Employment Levels by Gender and Country (Working Age: 15-64)") %>%
  hc_subtitle(text = "Interactive Year Selection | Including EA20 and EU27_2020") %>%
  hc_xAxis(
    categories = levels(summarized_data$geo),
    title = list(text = "Country"),
    labels = list(
      rotation = 45,
      style = list(fontSize = "8px")
    )
  ) %>%
  hc_yAxis(
    title = list(text = "Total Employment"),
    labels = list(format = "{value:,}")
  ) %>%
  hc_plotOptions(column = list(grouping = TRUE)) %>%
  hc_tooltip(shared = TRUE, pointFormat = "<b>{series.name}</b>: {point.y:,.0f}<br>")

# Add series for each year
for (yr in years) {
  year_data <- summarized_data %>% filter(year == yr)
  
  hchart <- hchart %>%
    hc_add_series(
      data = year_data %>% filter(sex == "M") %>% arrange(geo) %>% pull(total_employment),
      name = paste0("Male - ", yr),
      color = "#003399",
      visible = ifelse(yr == min(years), TRUE, FALSE)
    ) %>%
    hc_add_series(
      data = year_data %>% filter(sex == "F") %>% arrange(geo) %>% pull(total_employment),
      name = paste0("Female - ", yr),
      color = "#FFCC00",
      visible = ifelse(yr == min(years), TRUE, FALSE)
    )
}

# Add year selection dropdown and exporting options with a toggleable legend
hchart <- hchart %>%
  hc_exporting(enabled = TRUE) %>%
  hc_legend(enabled = TRUE) %>%
  hc_chart(events = list(
    load = JS("
      function() {
        var chart = this;
        var legendVisible = true;
        var button = chart.renderer.button('Toggle Legend', 10, 10)
          .on('click', function() {
            legendVisible = !legendVisible;
            chart.update({
              legend: { enabled: legendVisible },
              xAxis: {
                labels: {
                  style: {
                    fontSize: '8px'
                  },
                  rotation: 45
                }
              }
            });
            // Adjust the chart margins dynamically
            if (!legendVisible) {
              chart.update({
                chart: { marginRight: 50 }
              });
            } else {
              chart.update({
                chart: { marginRight: 150 }
              });
            }
          })
          .add();
      }")
  ))

# Display the chart
hchart
```

**Insert Comment here**

------------------------------------------------------------------------

# Employment's comparison by Country and Citizienship Category

A comparative analysis eas conducted to explore how employment levels
vary between different countries for specific citizenship categories.
`EU27_2020` and `EA20` have been excluded because they would make the
individual countries less comparable between eachother.

For this analysis a subset was created from the initial dataset,
including just the Working Population (`Y15-64`) and making no
differentiation for genre (`T`). Some specific years were selected: -
**2000**: Represents the early 2000s before major EU enlargements. -
**2007**: Captures the effects after major EU enlargements. - **2010**:
Captures employment trends post-2008 financial crisis. - **2023**:
Reflects the latest available data, including post-COVID-19 recovery.

Heatmaps were generated for each of these years and two comparisons were
made.

## Heatmaps Comparison - before and after the EU enlargements (2000 and 2007)

Heatmaps for 2000 and 2007 were generated and compared to capture the
differences in employment levels across countries **before** and
**after** the **EU enlargements**.

```{r, message=FALSE, warning=FALSE, out.width='100%'}
  # Filter data for selected years and only 'Total' for sex and working-age population (15-64)
  selected_years <- c("2000-01-01", "2007-01-01")
  filtered_data <- data %>%
    filter(
      TIME_PERIOD %in% as.Date(selected_years),
      sex == "T",
      age == "Y15-64",
      !geo %in% c("EU27_2020", "EA20")  # Exclude "EU27_2020" and "EA20"
    )
  
  # Group employment by country, citizenship, and year
  country_citizen_summary <- filtered_data %>%
    group_by(geo, citizen, TIME_PERIOD) %>%
    summarise(total_employment = sum(values, na.rm = TRUE))
  
  # Function to create a plotly heatmap for a specific year
  create_plotly_heatmap <- function(year, showlegend = TRUE) {
    df <- country_citizen_summary %>% filter(TIME_PERIOD == as.Date(year))
    
    # Multiply total_employment by 1,000 to reflect real values
    df <- df %>% mutate(total_employment_real = total_employment * 1000)
    
    plot_ly(
      data = df,
      x = ~geo,
      y = ~citizen,
      z = ~total_employment_real,
      type = "heatmap",
      colors = colorRamp(c("#f0f9ff", "#084594")),
      colorbar = list(
        title = "<b>Employment</b>",
        tickfont = list(size = 12),
        titlefont = list(size = 14, family = "Arial")
      ),
      showscale = showlegend,   # Control legend display
      zmin = 0, zmax = 50000000,   # Set consistent color scale limits
      
      # Custom hover text
        hovertemplate = paste(
          "<b>Country:</b> %{x}<br>",
          "<b>Citizenship:</b> %{y}<br>",
          "<b>Total Employment:</b> %{z}<br>",
          "<extra></extra>"  # Removes default trace info
        )
    ) %>%
      layout(
        title = list(
          #text = paste("<b>Employment by Country and Citizenship -", format(as.Date(year), "%Y"), "</b>"),
          font = list(size = 18, family = "Arial"),
          x = 0.5,  # Center the title
          xanchor = "center"
        ),
        xaxis = list(
          title = "<b>Country</b>",
          tickangle = 45,
          tickfont = list(size = 10),
          titlefont = list(size = 14, family = "Arial")
        ),
        yaxis = list(
          title = "<b>Citizenship Category</b>",
          tickfont = list(size = 10),
          titlefont = list(size = 14, family = "Arial")
        ),
        margin = list(t = 60, b = 60)  # Add padding to top and bottom margins
      )
  }
  
  # Create plotly heatmaps for each selected year
  interactive_heatmap_2000 <- create_plotly_heatmap("2000-01-01", showlegend = TRUE)
  interactive_heatmap_2007 <- create_plotly_heatmap("2007-01-01", showlegend = FALSE)

  # Combine the interactive heatmaps vertically
  combined_interactive_heatmaps <- subplot(
    interactive_heatmap_2000,
    interactive_heatmap_2007,
    nrows = 2,
    shareX = TRUE,
    shareY = TRUE,
    titleX = TRUE,
    titleY = TRUE
  ) %>%
    layout(
      annotations = list(
        list(
          text = "<b>Year: 2000 (before EU enlargements)</b>",
          x = 0.5,
          y = 1.06,
          xref = "paper",
          yref = "paper",
          showarrow = FALSE,
          font = list(size = 11, family = "Arial")
        ),
        list(
          text = "<b>Year: 2007 (after EU enlargements)</b>",
          x = 0.5,
          y = 0.50,
          xref = "paper",
          yref = "paper",
          showarrow = FALSE,
          font = list(size = 11, family = "Arial")
        )
      ),
      title = "<b>Employment Trends by Citizenship Category and Country (2000 vs 2007)</b>",
      margin = list(l = 100, r = 50, t = 80, b = 100),
      titlefont = list(size = 20, family = "Arial", color = "black")
    )
  
  # Display the combined interactive heatmaps
  combined_interactive_heatmaps
```

The heatmaps for the years **2000** and **2007** show notable variations
in employment levels across different **EU countries** and **citizenship
categories**. The color intensity represents employment levels, with
**darker blue** indicating higher employment.

### Year: 2000 (Before EU Enlargements)

-   **Employment for nationals** (`NAT`) dominates in most countries,
    reflecting a labor market primarily consisting of native citizens.
-   **Lower employment levels for foreign-born categories**
    (`EU27_2020_FOR` and `NEU27_2020_FOR`) are observed, as the European
    Union had not yet expanded to include countries from Eastern Europe
    (the 2004 and 2007 enlargements).
-   The employment distribution is more concentrated in **Western
    European countries** like **Germany** (`DE`), **France** (`FR`), and
    the **UK**.

### Year: 2007 (After EU Enlargements)

-   **Noticeable increase in employment for foreign-born categories**
    (`EU27_2020_FOR` and `NEU27_2020_FOR`), reflecting the impact of the
    **2004 EU enlargement** (when 10 new countries joined the EU) and
    anticipation of the **2007 enlargement** (including Bulgaria and
    Romania).
-   Employment growth in countries like **Germany** (`DE`), the **UK**,
    and **Spain** (`ES`) suggests increased **labor migration** and
    **workforce integration** of foreign-born citizens.
-   Some **Eastern European countries** begin to show increased
    employment for nationals and foreign-born categories, indicating the
    effect of **free movement of workers** within the EU.

### Key Insights

-   **Impact of EU Enlargement**:\
    The period between **2000 and 2007** shows the effects of **EU
    enlargement**, with a rise in employment for foreign-born categories
    due to increased migration and labor mobility within the EU.

-   **Country-Specific Patterns**:

    -   **Germany** (`DE`) and **France** (`FR`) maintain high
        employment levels across both years, reflecting their strong
        economies and capacity to integrate foreign workers.\
    -   **Spain** (`ES`) shows increased employment for foreign-born
        categories, consistent with the **construction boom** and
        economic growth before the **2008 financial crisis**.

-   **Labor Market Integration**:\
    The increase in employment for foreign-born categories highlights
    the **integration of migrants** into the workforce and the
    importance of policies supporting labor mobility and inclusion.

## Heatmaps Comparison - After 2 Important Crisis (2010 and 2023)

Heatmaps for 2010 and 2023 were compared to capture the differences in
employment levels across countries after 2 important crisis: **2008
Financial Crisis** and **COVID-19 Pandemic**.

```{r, message=FALSE, warning=FALSE, out.width='100%'}
  # Filter data for selected years and only 'Total' for sex and working-age population (15-64)
  selected_years <- c("2010-01-01", "2023-01-01")
  filtered_data <- data %>%
    filter(
      TIME_PERIOD %in% as.Date(selected_years),
      sex == "T",
      age == "Y15-64",
      !geo %in% c("EU27_2020", "EA20")  # Exclude "EU27_2020" and "EA20"
    )
  
  # Group employment by country, citizenship, and year
  country_citizen_summary <- filtered_data %>%
    group_by(geo, citizen, TIME_PERIOD) %>%
    summarise(total_employment = sum(values, na.rm = TRUE))
  
  # Function to create a plotly heatmap for a specific year
  create_plotly_heatmap <- function(year, showlegend = TRUE) {
    df <- country_citizen_summary %>% filter(TIME_PERIOD == as.Date(year))
    
    # Multiply total_employment by 1,000 to reflect real values
    df <- df %>% mutate(total_employment_real = total_employment * 1000)
    
    plot_ly(
      data = df,
      x = ~geo,
      y = ~citizen,
      z = ~total_employment_real,
      type = "heatmap",
      colors = colorRamp(c("#f0f9ff", "#084594")),
      colorbar = list(
        title = "<b>Employment</b>",
        tickfont = list(size = 12),
        titlefont = list(size = 14, family = "Arial")
      ),
      showscale = showlegend,   # Control legend display
      zmin = 0, zmax = 50000000,   # Set consistent color scale limits
      
      # Custom hover text
        hovertemplate = paste(
          "<b>Country:</b> %{x}<br>",
          "<b>Citizenship:</b> %{y}<br>",
          "<b>Total Employment:</b> %{z}<br>",
          "<extra></extra>"  # Removes default trace info
        )
    ) %>%
      layout(
        title = list(
          #text = paste("<b>Employment by Country and Citizenship -", format(as.Date(year), "%Y"), "</b>"),
          font = list(size = 18, family = "Arial"),
          x = 0.5,  # Center the title
          xanchor = "center"
        ),
        xaxis = list(
          title = "<b>Country</b>",
          tickangle = 45,
          tickfont = list(size = 10),
          titlefont = list(size = 14, family = "Arial")
        ),
        yaxis = list(
          title = "<b>Citizenship Category</b>",
          tickfont = list(size = 10),
          titlefont = list(size = 14, family = "Arial")
        ),
        margin = list(t = 60, b = 60)  # Add padding to top and bottom margins
      )
  }
  
  # Create plotly heatmaps for each selected year
  interactive_heatmap_2010 <- create_plotly_heatmap("2010-01-01", showlegend = TRUE)
  interactive_heatmap_2023 <- create_plotly_heatmap("2023-01-01", showlegend = FALSE)

  # Combine the interactive heatmaps vertically
  combined_interactive_heatmaps <- subplot(
    interactive_heatmap_2010,
    interactive_heatmap_2023,
    nrows = 2,
    shareX = TRUE,
    shareY = TRUE,
    titleX = TRUE,
    titleY = TRUE
  ) %>%
    layout(
      annotations = list(
        list(
          text = "<b>Year: 2010 (after 2008 Financial Crisis)</b>",
          x = 0.5,
          y = 1.06,
          xref = "paper",
          yref = "paper",
          showarrow = FALSE,
          font = list(size = 11, family = "Arial")
        ),
        list(
          text = "<b>Year: 2023 (after COVID-19 Pandemic)</b>",
          x = 0.5,
          y = 0.50,
          xref = "paper",
          yref = "paper",
          showarrow = FALSE,
          font = list(size = 11, family = "Arial")
        )
      ),
      title = "<b>Employment Trends by Citizenship Category and Country (2010 vs 2023)</b>",
      margin = list(l = 100, r = 50, t = 80, b = 100),
      titlefont = list(size = 20, family = "Arial", color = "black")
    )
  
  # Display the combined interactive heatmaps
  combined_interactive_heatmaps
```

The heatmaps for the years **2010** and **2023** show variations in
employment levels across different **EU countries** and **citizenship
categories**. The color intensity represents employment levels, with
**darker blue** indicating higher employment.

### Year: 2010 (After 2008 Financial Crisis)

-   **Employment for nationals** (`NAT`) remains prominent, but declines
    are observed in some countries due to the impact of the **2008
    financial crisis**.
-   **Lower employment levels for foreign-born categories**
    (`EU27_2020_FOR` and `NEU27_2020_FOR`), reflecting the economic
    downturn's disproportionate impact on migrant workers.
-   Employment growth is still concentrated in countries like
    **Germany** (`DE`) and **France** (`FR`), which were more resilient
    to the crisis.
-   **Southern European countries** such as **Spain** (`ES`) and
    **Italy** (`IT`) show noticeable declines, consistent with the
    severe economic challenges they faced during this period.

### Year: 2023 (After COVID-19 Pandemic)

-   **Recovery in employment levels** for both nationals (`NAT`) and
    foreign-born categories (`EU27_2020_FOR` and `NEU27_2020_FOR`),
    indicating the labor market rebound after the **COVID-19 pandemic**.
-   Significant employment growth for foreign-born categories in
    countries like **Germany** (`DE`), **France** (`FR`), and **Spain**
    (`ES`), suggesting a return to pre-pandemic trends and increased
    workforce integration.
-   **Eastern European countries** show varying employment levels,
    reflecting ongoing economic adjustments and workforce mobility
    within the EU.
-   **Missing Values** appear for some Eastern countries (e.g.Ukraine
    `UK`, Turkey `TR`, North Macedonia `MK`), probably due to the
    ongoing conflict between Russia and Ukraine.

### Key Insights

-   **Impact of Economic Shocks**:\
    The period between **2010 and 2023** highlights the effects of two
    major economic shocks:
    -   The **2008 financial crisis** led to reduced employment,
        particularly for foreign-born categories.
    -   The **COVID-19 pandemic** caused temporary disruptions but was
        followed by a recovery in 2023.
-   **Country-Specific Patterns**:
    -   **Germany** (`DE`) and **France** (`FR`) continue to show high
        employment levels, demonstrating economic resilience and
        effective labor policies.
    -   **Spain** (`ES`) shows employment fluctuations, with a decline
        post-2008 and recovery post-pandemic.
-   **Labor Market Integration**:\
    The rebound in employment for foreign-born categories in **2023**
    underscores the importance of **migration policies** and **workforce
    integration programs** in sustaining labor markets.
