---
title: "Comparative Analysis of Dimensionality Reduction Techniques on Employment Data (2021-2023): PCA, MDS, t-SNE and UMAP"
author: "Davide Bittelli"
date: "December 8, 2024"
output: html_document
---

# Introduction

This report presents a **Principal Component Analysis (PCA)** of employment data from Eurostat. The goal is to:

1. **Identify patterns and relationships** between countries based on employment statistics for different citizen categories.
2. **Reduce dimensionality** for easier visualization and interpretation.
3. **Determine the contributions** of each variable (citizen category) to the principal components.
4. **Visualize countries' similarities and differences** in employment patterns.

---

# Libraries and Data Loading

Loading the necessary libraries for data manipulation, analysis, and visualization.

```{r load-libraries, message=FALSE}
# Load necessary libraries
library(tidyverse)     # For data manipulation and visualization
library(eurostat)      # For downloading Eurostat data
library(FactoMineR)    # For performing PCA
library(factoextra)    # For visualizing PCA results
library(plotly)        # For interactive plots
```

---

# Data Preparation

I Downloaded the `lfsa_egan` dataset and filtered it for: <br>
  - **Years**: 2021 to 2023 <br>
  - **Age Group**: 20-64 years (working-age population) <br>
  - **Sex**: Both sexes (`T`) <br>
  - **Unit**: Employment measured in thousands (`THS_PER`) <br>

```{r setup, message=FALSE, warning=FALSE}
# Download the 'lfsa_egan' dataset from Eurostat
data <- get_eurostat("lfsa_egan")

# Filter for the years 2021-2023, age 20-64, both sexes, and unit in thousands
data_filtered <- data %>%
  filter(TIME_PERIOD >= as.Date("2021-01-01") & TIME_PERIOD <= as.Date("2023-12-31"),
         age == "Y20-64", 
         sex == "T",
         unit == "THS_PER") %>%
  drop_na(values)  # Remove rows with missing values
```
I Inspected the unique values in the citizen and geo columns to get a better understanding of the data.

```{r}
unique(data$citizen)
unique(data$geo)

```
I also created a data frame that maps each country code and categories to its corresponding full name.

```{r}
# Create a named vector mapping country codes and categories to their full names
country_names <- c(
  AT = "Austria", BE = "Belgium", CH = "Switzerland", CY = "Cyprus", CZ = "Czech Republic",
  DE = "Germany", DK = "Denmark", EA20 = "Euro Area 20", EE = "Estonia", EL = "Greece",
  ES = "Spain", EU27_2020 = "European Union 27", FI = "Finland", FR = "France", HU = "Hungary",
  IE = "Ireland", IS = "Iceland", IT = "Italy", LU = "Luxembourg", ME = "Montenegro",
  MT = "Malta", NL = "Netherlands", NO = "Norway", PT = "Portugal", RS = "Serbia",
  SE = "Sweden", SI = "Slovenia", SK = "Slovakia", UK = "United Kingdom", BG = "Bulgaria",
  HR = "Croatia", LT = "Lithuania", LV = "Latvia", MK = "North Macedonia", PL = "Poland",
  RO = "Romania", BA = "Bosnia and Herzegovina", TR = "Turkey",
  
  # Categories
  EU27_2020_FOR = "EU27 Foreign Born", FOR = "Foreign Born", NAT = "National", 
  NEU27_2020_FOR = "Non-EU27 Foreign Born", NRP = "Non-Resident Population", 
  STLS = "Short-Term Labour Supply", TOTAL = "Total"
)


```

---

# Data Transformation

I decided to transform the dataset into a wide format with countries as rows and citizen categories as columns. Then I filled the missing values with 0.

```{r}
pca_data <- data_filtered %>%
  pivot_wider(names_from = citizen, values_from = values, values_fill = 0)
```

I prepared the data for PCA by removing non-numeric columns.

```{r}
pca_input <- pca_data %>% select(-geo, -age, -sex, -unit, -freq, -TIME_PERIOD)
```

---

# PCA Execution

I standardized the data and perform PCA.

```{r}
# Standardize the data
pca_data_scaled <- scale(pca_input)

# Perform PCA
pca_result <- PCA(pca_data_scaled, graph = FALSE)
```

---

# Results and Visualizations

## Scree Plot

The scree plot shows the percentage of variance explained by each principal component.

```{r}
fviz_eig(pca_result)
pca_result$eig
```

**Interpretation**: <br>
- The **first principal component** (Dim.1) explains **83.48%** of the variance. <br>
- The **second principal component** (Dim.2) explains **14.41%**. <br>
- **Together**, Dim.1 and Dim.2 account for **97.89%** of the total variance.

---

## Variable Contributions

### Contributions to the First Principal Component (Dim.1)

```{r}
fviz_contrib(pca_result, choice = "var", axes = 1)
```

**Interpretation**: <br>
The variables `EU27_2020_FOR`, `FOR`, `NAT`, `NEU27_2020_FOR`, and `TOTAL` contribute almost equally (**~17%**) to the first principal component. This component captures general employment trends across these categories.

### Contributions to the Second Principal Component (Dim.2)
```{r}
fviz_contrib(pca_result, choice = "var", axes = 2)
```

**Interpretation**: <br>
The NRP (Non-Resident Population) category dominates the second component with **98.82%** contribution, indicating that Dim.2 captures variations specific to this category.

### Contributions to the First Two Components
```{r}
fviz_contrib(pca_result, choice = "var", axes = 1:2)
```

---

## Individual Plot (Countries)

Visualize the countries in the PCA space, colored by their country codes.

```{r pca-plot, warning=FALSE, message = FALSE}
# Replace country codes with full names in pca_data$geo
pca_data$geo <- country_names[pca_data$geo]

# Convert geo to a factor
pca_data$geo <- as.factor(pca_data$geo)

# Create the PCA plot with country names
pca_plot <- fviz_pca_ind(
  pca_result,
  label = "none",
  habillage = pca_data$geo,   # This now contains the full country names
  addEllipses = TRUE
) +
  labs(
    title = "PCA of Employment Data by Country (2021-2023) - Interactive Plot",
    x = "Principal Component 1 (83.5% Variance Explained)",
    y = "Principal Component 2 (14.4% Variance Explained)",
  )

# Convert the plot to an interactive plot
interactive_plot <- ggplotly(pca_plot)

# Manually update the legend labels to remove ",1"
for (i in seq_along(interactive_plot$x$data)) {
  if (!is.null(interactive_plot$x$data[[i]]$name)) {
    interactive_plot$x$data[[i]]$name <- gsub(",1", "", interactive_plot$x$data[[i]]$name)
  }
}

# Display the interactive plot
interactive_plot
```

Countries are grouped based on their employment patterns. By looking at the plot I identified 4 macro-clusters: <br>
1. **Turkey** on the top <br>
2. **Euro Area 20** and **European Union 27** on the right <br>
3. **Germany** in the middle <br>
4. The remaining groups on the lower-left <br>
By taking a closer look at the group on the lower-left (draw a rectangle on the plot to zoom) I saw that **Italy**, **France** and **Spain** were plotted closer to eachother and further away from the other EU-countries. This may show that these countries share a similar employment distributions.

---

# Interpretation of Results

## Key Findings

- Dim.1 captures general employment trends across various citizen categories (`EU27_2020_FOR`, `FOR`, `NAT`, etc.).
- Dim.2 highlights the Non-Resident Population (`NRP`) as a distinct group.
- The first two components explain **97.89%** of the total variance, making them highly informative.

## Country Clusters

- The PCA plot shows clusters of countries with similar employment distributions.
- Differences between clusters can inform targeted employment policies or further investigations.

---

# Conclusions

This PCA analysis helps to:

  - Identify employment patterns across different citizen categories.
  - Understand the influence of specific categories like the Non-Resident Population (`NRP`).
  - Visualize similarities and differences between countries.
  - These insights can inform policies and highlight areas of interest for further analysis.
  
---

# References
**Eurostat Dataset**: `lfsa_egan`

